---
title: detectPVC User Guide
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{detectPVC User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8](inputenc)
---

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=9, fig.height=4.5,
               dev.args=list(pointsize=16))
```


detectPVC is an R package to detect premature ventricular complexes (PVCs) in data from a [Polar
H10](https://www.polar.com/us-en/sensors/h10-heart-rate-sensor) chest-strap heart rate sensor.

We have used the [ECGLogger app](https://www.ecglogger.com/) on an iPhone
to extract the Polar H10 data as a CSV file (or really a series of CSV
files, in one hour blocks), and the [rsleep](https://rsleep.org/)
package to detect R peaks in the ECG signal.


### Usage

The library comes with a ~40 sec sample data set, `h10`.

```{r}
library(detectPVC)
data(h10)
```

Convert the included times (which are time stamps from a Polar H10, in
1e-9 seconds) to a standard date-time values.

```{r}
h10$datetime <- convert_timestamp(h10$time)
```


Use `detect_peaks()` to detect R peaks in the ECG trace.

```{r}
peaks <- detect_peaks(h10$ecg)
```

Plot the data, and add points at the peaks. The function
`plot_ecg()` is a base-graphics-based plotting function to mimic a
traditional ECG trace.

```{r}
plot_ecg(h10$datetime, h10$ecg)
points(h10$datetime[peaks], h10$ecg[peaks], pch=16, col="slateblue")
```

Use `calc_peak_stats()` to calculate some statistics about each peak.

```{r}
peak_stats <- calc_peak_stats(peaks, h10$ecg)
```

The statistics `pmin` and `leftRR` seem particularly good for
identifying the PVCs.

```{r}
par(las=1, mar=c(4.1, 4.1, 0.6, 0.6))
plot(peak_stats$pmin, peak_stats$leftRR, xlab="peak min", ylab="left RR")
```

Label the inferred PVCs with pink dots, and the others with green
dots.

```{r}
plot_ecg(h10$datetime, h10$ecg)
pvc <- (peak_stats$pmin > -0.45)
points(h10$datetime[peaks[pvc]], peak_stats$pmax[pvc], pch=16, col="violetred")
points(h10$datetime[peaks[!pvc]], peak_stats$pmax[!pvc], pch=16, col="green3")
```

In this 40 second window, there are `r sum(pvc)` PVCs in `r length(pvc)` total beats.

---
