---
title: detectPVC User Guide
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{detectPVC User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8](inputenc)
---

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=9, fig.height=4.5,
               dev.args=list(pointsize=16))
```


detectPVC is an R package to detect premature ventricular complexes (PVCs) in data from a [Polar
H10](https://www.polar.com/us-en/sensors/h10-heart-rate-sensor) chest-strap heart rate sensor.

We have used the [ECGLogger app](https://www.ecglogger.com/) on an iPhone
to extract the Polar H10 data as a CSV file (or really a series of CSV
files, in one hour blocks), and the [rsleep](https://rsleep.org/)
package to detect R peaks in the ECG signal.


### Usage

The library comes with a ~10 min sample data set, `polar_h10`.

```{r load_data}
library(detectPVC)
data(polar_h10)
```

Convert the included times (which are time stamps from a Polar H10, in
nanoseconds) to a standard date-time values.

```{r convert_timestamp}
polar_h10$datetime <- convert_timestamp(polar_h10$time)
```


Use `detect_peaks()` to detect R peaks in the ECG trace.

```{r detect_peaks}
peaks <- detect_peaks(polar_h10$time, polar_h10$ecg)
```

Plot the first 20 seconds of the data, and add points above the peaks. The function
`plot_ecg()` is a base-graphics-based plotting function with light
gray grid lines.

```{r plot_ecg}
v <- 1:(130*20)
plot_ecg(polar_h10$datetime[v], polar_h10$ecg[v])
points(polar_h10$datetime[peaks], polar_h10$ecg[peaks], pch=16, col="slateblue")
```

Use `calc_peak_stats()` to calculate some statistics about each peak.

```{r calc_peak_stats}
peak_stats <- calc_peak_stats(polar_h10$time, polar_h10$ecg, peaks)
```

The simplest rule for classifying PVCs is to take `RStime > 50`.
The statistic `RStime` is the time between the R and S peaks in milliseconds.

```{r classify_pvc}
pvc <- (peak_stats$RStime > 50)
```

Label the inferred PVCs with pink dots, and the others with green
dots.

```{r plot_ecg_again}
plot_ecg(polar_h10$datetime[v], polar_h10$ecg[v])
points(polar_h10$datetime[peaks], polar_h10$ecg[peaks], pch=16, col=c("green3", "violetred")[pvc+1])
```

In this 20 second window, there are `r sum(pvc[peaks %in% v])` PVCs in
`r sum(peaks %in% v)` total beats.

---
